Causal Compatibility Inequalities Admitting of Quantum Violations in the Triangle Structure

### 2017-03-08 documents/TriangleScenarioArXiv/ts.tex
scenario -> structure, capitalized 'Triangle' in 'triangle scenario

\begin{document}
    \title{Causal Compatibility Inequalities Admitting of Quantum Violations in the Triangle Scenario}
    \author{Thomas C. Fraser}
    \email{tcfraser@tcfraser.com}
    \affiliation{Perimeter Institute for Theoretical Physics, Waterloo, Ontario, Canada \\ University of Waterloo, Waterloo, Ontario, Canada}
    % \author{Elie Wolfe}
    % \email{ewolfe@perimeter@institute.ca}
    % \affiliation{Perimeter Institute for Theoretical Physics, Waterloo, Ontario, Canada}
    \date{\today}
    \begin{abstract}
        It has long been recognized that certain quantum correlations are often incompatible with particular assumption about classical causal structure. Such correlations are sometimes termed nonlocal, although it would be more accurate to say simply noncassical, as the presence of correlations certifies the genuinely nonclassical nature of a causal setup in a device-independent fashion.  Whenever a  quantum correlation violates a causal compatibility inequality (e.g. a Bell inequality) then the correlation is witnessed as nonclassical. Any constraint which is satisfied by all correlations which \emph{are} classically compatible with a given causal scenario defined a causal compatibility criterion. Such criteria were recently were recently derived for the triangle scenario [\href[pdfnewwindow]{http://arxiv.org/abs/1609.00672}{arXiv:1609.00672}] in the form of polynomial inequalities, begging the question: do any of those inequalities admit violation by quantum correlations? Numerical investigation suggests not, and we further conjecture that the set of correlations admitted by the classical triangle triangle scenario is equivalent to the set of correlations admitted by its quantum generalization whenever the three observable variables are binary. Our main contribution inn this work, however, is the derivation of new causal compatibility inequalities for the triangle scenario which \emph{do} admit quantum violation. We conclude by considering the possibility of quantum resources potentially qualitatively different from those known previously.
    \end{abstract}
    \maketitle
    \tableofcontents
    \setlength{\parskip}{1em}
    \clearpage

@@

    From a foundational prospective, the most robust demonstrations of quantum phenemena with no classical emulation have involved Bell inequalities~\cite{Bell_1964}. Originally, Bell inequalities were derived as a way to show that no hidden variable theory could ever account for quantum mechanics; in this sense Bell inequalities are a response to the famous EPR paradox~\cite{EPR_Orig}. The enumeration of Bell inequalities has since become a widespread systematic way of demonstration the \textit{nonclassicality} of a given observation. More recently it has been appreciated that Bell inequalities can understood as consequences of \term{causal inference}~\cite{Wood_2012}. Causal inference is concerned with classifying observations into those which can and cannot be explained by a hypothesized causal structure. The abstract nature of causal inference is responsible for its presence in numerous scientific fields including machine learning and biology~\cite{Pearl_2009,Pearl_2009_tr}. Causal compatability inequalities, such as Bell of Instrumental inequalities, charechterize the space of observations that are compatible with a hypothesized causal structure, albeit the charechterization offered by practically deriveable inequalities often only an approximation. To derive the traditional Bell inequalities from causal inference one starts with a (classical) causal structure known as the Bell scenario, as depicted in \cref{fig:bell_scenario}. The fundamental Bell scenario involves noncommunicating parties making measurements on some hidden shared resource $\la$, where the measuremement outcomes ($A, B$) are presumed to be stochastic functions of the the local choices of measurement settings ($S_A, S_B$) and the shared resource $\la$. Quantum nonclassicality in the Bell scenario has been thoroughly studied since Bell's original work~\cite{Brunner_2013}. More complex scenarios, however, such as the correlation scenarios proposed by Fritz~\cite{Fritz_2012,Fritz_2014}, are much less understood. Here we investigate one particular causal scenario: the triangle scenario (\cref{fig:triangle_scenario}).

    The triangle scenario (\cref{fig:triangle_scenario}) is a causal structure composed of $3$ parties labeled $A, B, C$ arranged in a triangular configuration while pair-wise sharing hidden/latent variables $X, Y, Z$. It has been extensively studied previously (see~\cite[Fig. 1]{Steudel_2010},~\cite[Fig. 6]{Chaves_2014},~\cite[Fig. 8]{Branciard_2012},~\cite[Fig. 8, App. E]{Henson_2014},~\cite[Fig. 3]{Fritz_2012},~\cite[Fig. 1]{Inflation}, $\ldots$). An overview of some milestone results is provided in \cref{sec:triangle_scenario}. Identifying causal compatibility inequalities for this configuration has been seen as particularly challenging~\citet{Branciard_2012}. Further identifying causal compatibility inequalities of such high resolution that such inequalities can be violated by quantum-accessible distributions has remained out-of-reach for the triangle scenario.

    This work is the first, therefore, to find causal compatibility inequalities for the triangle scenario that are able to be violated by quantum-accessible distributions. This accomplishment was made possible through the combination of two previous developments. The first, an insight made by \citet{Fritz_2012} regarding the ability to re-interpret the Bell scenario as a portion of the triangle scenario. The second, a new framework for solving causal inference problems developed by Wolfe \emph{et al.}~\cite{Inflation} called \term{The Inflation Technique}. Ultimately, this work serves as a validation that the Inflation Technique is sensitive enough to offer new insights into quantum nonclassicality. Moreover, these new inequalities offer an avenue for recognizing previously unknown forms of nonclassicality. The authors attempts to find such novel resources were met with only partial success, suggesting clear refinements for future exploration.

    The manuscript is organized into the following sections: \Cref{sec:causal_compatibility} recalls important notions from causal inference theory and sets up the notation to be used. \Cref{sec:triangle_scenario} discusses the triangle scenario and provides an overview of existing research; identifying its stark differences from the Bell scenario and motivating why the triangle scenario was chosen as a case study. \Cref{sec:fritz_distribution} defines and discusses a singularly-quantum correlation first concieved of by \citet{Fritz_2012}, which we term the \term{Fritz distribution}. The Fritz distribution was proven to be non-classical without the use of inequalities --- this report fills in the desired missing inequalities. \Cref{sec:inflation_technique} briefly summaries the Inflation Technique in the specific context of this work. Although the summary presented here is designed to be self-standing, a much more pedagogical introduction is offered by the original work~\cite{Inflation}. \Cref{sec:deriving_inequalities} demonstrates how the inflation technique can be used to derive causal compatibility inequalities for the triangle scenario that are violated by quantum-accessible distributions. A sample of these inequalities are presented at the end of \cref{sec:deriving_inequalities} as well as in \cref{sec:web_inequality,sec:symmetic_web_inequality}. Finally, \cref{sec:violations_noise} introduces the method for parameterizing the space of quantum-accessible distributions which we used to numerically optimize against our derived inequalities in search of stronger types of non-classicality. Additionally, the Fritz distribution is subjected to noise in order to measure the robustness of the derived inequalities.

@@

    A number of causal structures are depicted throughout this document (see \cref{fig:bell_scenario,fig:triangle_scenario} as well as \cref{fig:inflations}). As an example, the triangle scenario has $3$ observable nodes and $3$ latent nodes,
    \[ \nodes_{O} = \bc{A, B, C} \qquad \nodes_{L} = \bc{X, Y, Z} \]
    along with $6$ edges indicating that the latent nodes $\bc{X, Y, Z}$ cause and influence the observed nodes $\bc{A, B, C}$.
    \[ \edges = \bc{\bc{X \to A}, \bc{Y \to A}, \bc{Y \to B}, \bc{Z \to B}, \bc{Z \to C}, \bc{X \to C}} \]

@@

    \subsection{Inequalities}
    As defined in \cref{sec:compatibility}, determining compatibility is equivalent to determining if a set of causal parameters exists with the above-mentioned properties. In practice however, checking causal parameters systemically is intractable, especially when the cardinality (number of outcomes) of the latent variables is unspecified. Instead, \term{causal compatibility inequalities}\footnote{We refer to these inequalities as ``causal compatibility inequalities'' instead of simply Bell inequalities for two reasons. Firstly, ``Bell inequalities'' usually are associated specifically with the Bell scenario/Bell causal structure. Secondly, the inequalities derived in this work are fundamentally distinct from a typical Bell inequality in that these inequalities are \textit{polynomial} over $\prob^{\mscenario}$ instead of \textit{linear}.} are probabilistic inequalities defined over the marginal contexts $\prob[V] \in \prob^{\mscenario}$ within the marginal model. If a marginal model $\prob^{\mscenario}$ happens to \textit{violate} any causal compatibility inequality, then that marginal model is definitely \textit{incompatible}. In \cref{sec:deriving_inequalities}, a brief discussion of existing techniques for deriving these inequalities will take place.

    Unfortunately, a singular inequality can not prove that a given marginal model is \textit{compatible}, only \textit{incompatible}. Instead, a \textit{complete characterization} of compatibility would consist of a complete set of all valid causal compatibility inequalities; satisfaction of a complete set of inequalities does prove compatibility. Currently however, it is unknown how to obtain a complete characterization for any given causal structure, including the triangle scenario.

@@

    \section{Triangle Scenario}
    \label{sec:triangle_scenario}

    \begin{figure}
    \begin{center}
        \begin{minipage}[t]{.48\textwidth}
            \centering
            \scalebox{1.0}{\input{../../figures/causal_structures/bell_scenario}}
            \caption{The Bell scenario consisting of two observers $\p{A}, \p{B}$ together with measurement settings $S_{\p{A}}$ and $S_{\p{B}}$ respectively. The shared hidden variable is labeled $\la$.}
            \label{fig:bell_scenario}
        \end{minipage}\hspace{0.04\textwidth}%
        \begin{minipage}[t]{.48\textwidth}
            \centering
            \scalebox{1.0}{\input{../../figures/causal_structures/triangle_scenario}}
            \caption{The casual structure of the triangle scenario. Three variables $A,B,C$ are observable, while $X, Y, Z$ are latent variables.}
            \label{fig:triangle_scenario}
        \end{minipage}
    \end{center}
    \end{figure}
    Discovering quantum non-classicality in any causal structure is desirable, so why, in particular, is it important to do so in the triangle scenario? As was mentioned in \cref{sec:introduction,sec:causal_compatibility}, the \term{triangle scenario} (\cref{fig:triangle_scenario}) is a causal structure $\graph$ consisting of $3$ observable variables $A, B, C$ arranged in a triangular configuration while pair-wise sharing latent variables $X, Y, Z$.

    Following the definition of causal compatibility from \cref{sec:causal_compatibility}, a distribution $\prob[\nodes_{O}] = \prob[ABC]$ is compatible with the triangle scenario if and only if there exists a choice of causal parameters $\bc{\prob[A|X,Y],\prob[B|Y,Z],\prob[C|Z,X],\prob[X],\prob[Y],\prob[Z]}$ such that 
    %the joint distribution $\prob[\nodes] = \prob[ABCXYZ]$ is a product over the causal parameters,
    %\[ \prob[ABCXYZ] = \prob[A|X,Y]\prob[B|Y,Z]\prob[C|Z,X]\prob[X]\prob[Y]\prob[Z] \]
    %And also, the given distribution 
    $\prob[ABC]$ is a marginalization of $\prob[ABCXYZ]$ over $X, Y, Z$:
    \[ \prob[ABC] = \sum_{X,Y,Z}\prob[A|X,Y]\prob[B|Y,Z]\prob[C|Z,X]\prob[X]\prob[Y]\prob[Z] \eq \label{eq:triangle_compatibility} \]

    For certain causal structures, the conditional independence relations provided by $d$-separation relations are a sufficient characterization of compatibility. Recently, \citet{Henson_2014}, classified the triangle scenario as an \textit{interesting} causal structure in the sense that conditional independence relations \textit{do not} provide a sufficient characterization of non-classicality.\footnote{Note that the triangle scenario possesses \textit{zero} conditional independence relations over the observable nodes $A, B, C$.} Moreover, \citet{Inflation} concretely demonstrate that the space of classical distributions on the triangle scenario is non-convex, unlike the Bell scenario. It can be argued that convexity of the Bell scenario distributions is partially responsible for the wealth of knowledge that is known about the Bell scenario. It is also reasonable to speculate that quantum non-classicality in the triangle can be translated directly into a computational advantage for certain computational circuits~\cite{Terhal_2002}. Characterizing quantum non-classicality for interesting causal structures remains an unsolved problem, and the triangle scenario is a small, non-trivial case study.

    Additionally, \citet{Fritz_2012} demonstrated that the triangle scenario is the \textit{smallest} correlation scenario\footnote{Correlation scenarios differ from the Bell scenario in that measure setting variables are non-existent in correlation scenarios.} in which their exists incompatible, quantum distributions by explicitly mentioning one which we have elected to call the Fritz distribution (see \cref{sec:fritz_distribution}). Fritz's proof of non-classicality does not involve compatibility inequalities, but instead mentions that inequalities would be helpful for characterizing non-classical in the triangle scenario.
     \comment[E]{Revisions processed up to here thusfar.}

    Finally \citet{Inflation} proposed the Inflation Technique (summarized in \cref{sec:inflation_technique}), a novel tool for solving causal inference problems. Using the inflation technique, they were able to prove incompatibility between the triangle scenario and the $W$-type distribution; a quantum non-accessible distribution whose incompatibility is not witness-able by \textit{any} entropic inequalities or any other known constraints. Therefore as a an algorithmic framework, the Inflation Technique comes closer to characterizing non-classicality in the triangle scenario that all others.

    It is worth noting that a number of causal compatibility inequalities for the triangle scenario have been derived in previous works. For example, \citet{Steudel_2010} derived an inequality for a series of causal structures with $n \in \N$ variables where at most $c \in \N$ share a common ancestor. The triangle scenario is a special case with $c = 2$, $n = 3$. Moreover, \citet{Henson_2014} derive a family of entropic inequalities constraining the set of compatible correlations. Most impressively, \citet{Inflation} derive a family of polynomial, $2$-outcome, causal compatibility inequalities; $37$ non-trivial representatives are printed in~\cite{Inflation}.

    As a preliminary search for quantum incompatibility in the triangle scenario, we performed numerical optimizations using bipartite qubit density matrices and $2$-outcome POVM measurements against the compatibility inequalities printed in~\cite{Inflation} as well as the entropic inequalities presented by \citet{Henson_2014}. Unfortunately, none of these inequalities could be violated. These early results suggest, although do not prove, that quantum non-classical does not exist in the triangle scenario for \textit{two-outcome} measurements. We make \textit{no claim} that this is a universal truth and simply take it as motivation to explore incompatibility in triangle scenario for a degree of outcomes.

    In conclusion, the triangle scenario is a causal structure whose non-classicality remains an open problem. The triangle scenario is a desirable case study because there is a quantum distribution that is known to be incompatible, but no previously known inequalities were able to witness its incompatibility. This failure represents a gap in our understanding of quantum non-classicality, where recent developments (namely the Inflation Technique) have made partial progress. This document expands on our understanding of non-classicality in the triangle scenario and also our understanding on how to find inequalities using the Inflation Technique for interesting causal structures in general.

    \section{Fritz Distribution}
    \label{sec:fritz_distribution}
    \begin{figure}
    \begin{center}
        \scalebox{1.0}{\input{../../figures/causal_structures/triangle_scenario_operational_fritz_bell}}
        \caption{The triangle scenario re-imagined to mimic the Bell scenario. The measurement settings $S_{\p{A}},S_{\p{B}}$ are latent nodes unlike the Bell scenario (\cref{fig:bell_scenario}).}
        \label{fig:triangle_scenario_with_fritz_bell_embedded}
    \end{center}
    \end{figure}

    As was originally noticed by \citet{Fritz_2012}, it is possible to construct quantum distributions incompatible with the triangle scenario by utilizing quantum distributions incompatible with the familiar Bell scenario. To explain, imagine rearranging the triangle into the configuration depicted in \cref{fig:triangle_scenario_with_fritz_bell_embedded} and contrast it with the Bell scenario of \cref{fig:bell_scenario}. Evidently, under the correct relabeling, large portions of the Triangle Scenario resemble the Bell Scenario. The crucial distinction is that the measurement settings $S_{\p{A}}, S_{\p{B}}$ are \textit{observable variables} in the Bell scenario but become \textit{latent variables} in the triangle scenario. In order to embed distributions that are incompatible with the Bell scenario into the triangle scenario, the latent nodes $S_{\p{A}}, S_{\p{B}}$ need to be observable and independent of the shared latent node $\la$. This can be accomplished by having $C$ ``measure'' the measurement settings $S_{\p{A}}, S_{\p{B}}$ and announce them as an outcome. Consequently, any distribution over $\p{A}, \p{B}, S_{\p{A}}, S_{\p{B}}$ that is incompatible with Bell scenario is also incompatible with the triangle scenario provided that $C$ is \textit{perfectly correlated} with $S_{\p{A}}, S_{\p{B}}$. Understandably there are numerous ways that this can be accomplished, albeit we will focus on elucidating an exemplary method.

    The \term{Fritz distribution}, denoted $\prob[\fritz]$, is a quantum-accessible distribution known to be incompatible with the triangle scenario~\cite{Fritz_2012}. In the Fritz distribution, each of the variables $A,B,C$ have $4$ possible outcomes. Explicitly, $\prob[\fritz]$ can be written as:

@@

    Here the notation $\prob[\fritz][abc] = \prob[\p{ABC}][abc] = \prob[][\p A=a,\p B=b,\p C=c]$ is used as shorthand. The Fritz distribution is best visualized as a $4 \times 4 \times 4$ grid of possible outcomes as depicted in \cref{fig:fritz_distribution_visualized}. From this diagram, it can be seen that each of $\p{C}$'s outcomes restricts the possible outcomes for $\p A,\p B$ into a $2 \times 2$ block. If one writes the outcome labels $\bc{0,1,2,3}$ in binary $\bc{00,01,10,11}$, it can be seen that the left-hand bits for $\p A$ and $\p B$ (respectively denoted $\p A_l$, $\p B_l$) are fixed by the outcome of $\p C$. Pursuant to the embedding of~\cref{fig:triangle_scenario_with_fritz_bell_embedded} the left-hand bits emulate the measurement settings $S_{\p A}$ and $S_{\p B}$ and the right-hand bits emulate a two-outcome measurement performed by $A, B$ as they would in~\cref{fig:bell_scenario}. Specifically, $\p C$'s bits are perfectly correlated with the left-bits of $A,B$; $\p C_l = \p A_l$ and $\p C_r = \p B_l$.

>>>>

\begin{document}
    \title{Causal Compatibility Inequalities Admitting of Quantum Violations in the Triangle Structure}
    \author{Thomas C. Fraser}
    \email{tcfraser@tcfraser.com}
    \affiliation{Perimeter Institute for Theoretical Physics, Waterloo, Ontario, Canada \\ University of Waterloo, Waterloo, Ontario, Canada}
    % \author{Elie Wolfe}
    % \email{ewolfe@perimeter@institute.ca}
    % \affiliation{Perimeter Institute for Theoretical Physics, Waterloo, Ontario, Canada}
    \date{\today}
    \begin{abstract}
        It has long been recognized that certain quantum correlations are often incompatible with particular assumption about classical causal structure. Such correlations are sometimes termed nonlocal, although it would be more accurate to say simply noncassical, as the presence of correlations certifies the genuinely nonclassical nature of a causal setup in a device-independent fashion.  Whenever a  quantum correlation violates a causal compatibility inequality (e.g. a Bell inequality) then the correlation is witnessed as nonclassical. Any constraint which is satisfied by all correlations which \emph{are} classically compatible with a given causal scenario defined a causal compatibility criterion. Such criteria were recently were recently derived for the Triangle structure [\href[pdfnewwindow]{http://arxiv.org/abs/1609.00672}{arXiv:1609.00672}] in the form of polynomial inequalities, begging the question: do any of those inequalities admit violation by quantum correlations? Numerical investigation suggests not, and we further conjecture that the set of correlations admitted by the classical triangle Triangle structure is equivalent to the set of correlations admitted by its quantum generalization whenever the three observable variables are binary. Our main contribution inn this work, however, is the derivation of new causal compatibility inequalities for the Triangle structure which \emph{do} admit quantum violation. We conclude by considering the possibility of quantum resources potentially qualitatively different from those known previously.
    \end{abstract}
    \maketitle
    \tableofcontents
    \setlength{\parskip}{1em}
    \clearpage

@@

    From a foundational prospective, the most robust demonstrations of quantum phenemena with no classical emulation have involved Bell inequalities~\cite{Bell_1964}. Originally, Bell inequalities were derived as a way to show that no hidden variable theory could ever account for quantum mechanics; in this sense Bell inequalities are a response to the famous EPR paradox~\cite{EPR_Orig}. The enumeration of Bell inequalities has since become a widespread systematic way of demonstration the \textit{nonclassicality} of a given observation. More recently it has been appreciated that Bell inequalities can understood as consequences of \term{causal inference}~\cite{Wood_2012}. Causal inference is concerned with classifying observations into those which can and cannot be explained by a hypothesized causal structure. The abstract nature of causal inference is responsible for its presence in numerous scientific fields including machine learning and biology~\cite{Pearl_2009,Pearl_2009_tr}. Causal compatability inequalities, such as Bell of Instrumental inequalities, charechterize the space of observations that are compatible with a hypothesized causal structure, albeit the charechterization offered by practically deriveable inequalities often only an approximation. To derive the traditional Bell inequalities from causal inference one starts with a (classical) causal structure known as the Bell structure, as depicted in \cref{fig:bell_scenario}. The fundamental Bell structure involves noncommunicating parties making measurements on some hidden shared resource $\la$, where the measuremement outcomes ($A, B$) are presumed to be stochastic functions of the the local choices of measurement settings ($S_A, S_B$) and the shared resource $\la$. Quantum nonclassicality in the Bell structure has been thoroughly studied since Bell's original work~\cite{Brunner_2013}. More complex scenarios, however, such as the correlation scenarios proposed by Fritz~\cite{Fritz_2012,Fritz_2014}, are much less understood. Here we investigate one particular causal scenario: the Triangle structure (\cref{fig:triangle_structure}).

    The Triangle structure (\cref{fig:triangle_structure}) is a causal structure composed of $3$ parties labeled $A, B, C$ arranged in a triangular configuration while pair-wise sharing hidden/latent variables $X, Y, Z$. It has been extensively studied previously (see~\cite[Fig. 1]{Steudel_2010},~\cite[Fig. 6]{Chaves_2014},~\cite[Fig. 8]{Branciard_2012},~\cite[Fig. 8, App. E]{Henson_2014},~\cite[Fig. 3]{Fritz_2012},~\cite[Fig. 1]{Inflation}, $\ldots$). An overview of some milestone results is provided in \cref{sec:triangle_structure}. Identifying causal compatibility inequalities for this configuration has been seen as particularly challenging~\citet{Branciard_2012}. Further identifying causal compatibility inequalities of such high resolution that such inequalities can be violated by quantum-accessible distributions has remained out-of-reach for the Triangle structure.

    This work is the first, therefore, to find causal compatibility inequalities for the Triangle structure that are able to be violated by quantum-accessible distributions. This accomplishment was made possible through the combination of two previous developments. The first, an insight made by \citet{Fritz_2012} regarding the ability to re-interpret the Bell structure as a portion of the Triangle structure. The second, a new framework for solving causal inference problems developed by Wolfe \emph{et al.}~\cite{Inflation} called \term{The Inflation Technique}. Ultimately, this work serves as a validation that the Inflation Technique is sensitive enough to offer new insights into quantum nonclassicality. Moreover, these new inequalities offer an avenue for recognizing previously unknown forms of nonclassicality. The authors attempts to find such novel resources were met with only partial success, suggesting clear refinements for future exploration.

    The manuscript is organized into the following sections: \Cref{sec:causal_compatibility} recalls important notions from causal inference theory and sets up the notation to be used. \Cref{sec:triangle_structure} discusses the Triangle structure and provides an overview of existing research; identifying its stark differences from the Bell structure and motivating why the Triangle structure was chosen as a case study. \Cref{sec:fritz_distribution} defines and discusses a singularly-quantum correlation first concieved of by \citet{Fritz_2012}, which we term the \term{Fritz distribution}. The Fritz distribution was proven to be non-classical without the use of inequalities --- this report fills in the desired missing inequalities. \Cref{sec:inflation_technique} briefly summaries the Inflation Technique in the specific context of this work. Although the summary presented here is designed to be self-standing, a much more pedagogical introduction is offered by the original work~\cite{Inflation}. \Cref{sec:deriving_inequalities} demonstrates how the inflation technique can be used to derive causal compatibility inequalities for the Triangle structure that are violated by quantum-accessible distributions. A sample of these inequalities are presented at the end of \cref{sec:deriving_inequalities} as well as in \cref{sec:web_inequality,sec:symmetic_web_inequality}. Finally, \cref{sec:violations_noise} introduces the method for parameterizing the space of quantum-accessible distributions which we used to numerically optimize against our derived inequalities in search of stronger types of non-classicality. Additionally, the Fritz distribution is subjected to noise in order to measure the robustness of the derived inequalities.

@@

    A number of causal structures are depicted throughout this document (see \cref{fig:bell_scenario,fig:triangle_structure} as well as \cref{fig:inflations}). As an example, the Triangle structure has $3$ observable nodes and $3$ latent nodes,
    \[ \nodes_{O} = \bc{A, B, C} \qquad \nodes_{L} = \bc{X, Y, Z} \]
    along with $6$ edges indicating that the latent nodes $\bc{X, Y, Z}$ cause and influence the observed nodes $\bc{A, B, C}$.
    \[ \edges = \bc{\bc{X \to A}, \bc{Y \to A}, \bc{Y \to B}, \bc{Z \to B}, \bc{Z \to C}, \bc{X \to C}} \]

@@

    \subsection{Inequalities}
    As defined in \cref{sec:compatibility}, determining compatibility is equivalent to determining if a set of causal parameters exists with the above-mentioned properties. In practice however, checking causal parameters systemically is intractable, especially when the cardinality (number of outcomes) of the latent variables is unspecified. Instead, \term{causal compatibility inequalities}\footnote{We refer to these inequalities as ``causal compatibility inequalities'' instead of simply Bell inequalities for two reasons. Firstly, ``Bell inequalities'' usually are associated specifically with the Bell structure/Bell causal structure. Secondly, the inequalities derived in this work are fundamentally distinct from a typical Bell inequality in that these inequalities are \textit{polynomial} over $\prob^{\mscenario}$ instead of \textit{linear}.} are probabilistic inequalities defined over the marginal contexts $\prob[V] \in \prob^{\mscenario}$ within the marginal model. If a marginal model $\prob^{\mscenario}$ happens to \textit{violate} any causal compatibility inequality, then that marginal model is definitely \textit{incompatible}. In \cref{sec:deriving_inequalities}, a brief discussion of existing techniques for deriving these inequalities will take place.

    Unfortunately, a singular inequality can not prove that a given marginal model is \textit{compatible}, only \textit{incompatible}. Instead, a \textit{complete characterization} of compatibility would consist of a complete set of all valid causal compatibility inequalities; satisfaction of a complete set of inequalities does prove compatibility. Currently however, it is unknown how to obtain a complete characterization for any given causal structure, including the Triangle structure.

@@

    \section{Triangle Structure}
    \label{sec:triangle_structure}

    \begin{figure}
    \begin{center}
        \begin{minipage}[t]{.48\textwidth}
            \centering
            \scalebox{1.0}{\input{../../figures/causal_structures/bell_scenario}}
            \caption{The Bell structure consisting of two observers $\p{A}, \p{B}$ together with measurement settings $S_{\p{A}}$ and $S_{\p{B}}$ respectively. The shared hidden variable is labeled $\la$.}
            \label{fig:bell_scenario}
        \end{minipage}\hspace{0.04\textwidth}%
        \begin{minipage}[t]{.48\textwidth}
            \centering
            \scalebox{1.0}{\input{../../figures/causal_structures/triangle_structure}}
            \caption{The casual structure of the Triangle structure. Three variables $A,B,C$ are observable, while $X, Y, Z$ are latent variables.}
            \label{fig:triangle_structure}
        \end{minipage}
    \end{center}
    \end{figure}
    Discovering quantum non-classicality in any causal structure is desirable, so why, in particular, is it important to do so in the Triangle structure? As was mentioned in \cref{sec:introduction,sec:causal_compatibility}, the \term{Triangle structure} (\cref{fig:triangle_structure}) is a causal structure $\graph$ consisting of $3$ observable variables $A, B, C$ arranged in a triangular configuration while pair-wise sharing latent variables $X, Y, Z$.

    Following the definition of causal compatibility from \cref{sec:causal_compatibility}, a distribution $\prob[\nodes_{O}] = \prob[ABC]$ is compatible with the Triangle structure if and only if there exists a choice of causal parameters $\bc{\prob[A|X,Y],\prob[B|Y,Z],\prob[C|Z,X],\prob[X],\prob[Y],\prob[Z]}$ such that
    %the joint distribution $\prob[\nodes] = \prob[ABCXYZ]$ is a product over the causal parameters,
    %\[ \prob[ABCXYZ] = \prob[A|X,Y]\prob[B|Y,Z]\prob[C|Z,X]\prob[X]\prob[Y]\prob[Z] \]
    %And also, the given distribution
    $\prob[ABC]$ is a marginalization of $\prob[ABCXYZ]$ over $X, Y, Z$:
    \[ \prob[ABC] = \sum_{X,Y,Z}\prob[A|X,Y]\prob[B|Y,Z]\prob[C|Z,X]\prob[X]\prob[Y]\prob[Z] \eq \label{eq:triangle_compatibility} \]

    For certain causal structures, the conditional independence relations provided by $d$-separation relations are a sufficient characterization of compatibility. Recently, \citet{Henson_2014}, classified the Triangle structure as an \textit{interesting} causal structure in the sense that conditional independence relations \textit{do not} provide a sufficient characterization of non-classicality.\footnote{Note that the Triangle structure possesses \textit{zero} conditional independence relations over the observable nodes $A, B, C$.} Moreover, \citet{Inflation} concretely demonstrate that the space of classical distributions on the Triangle structure is non-convex, unlike the Bell structure. It can be argued that convexity of the Bell structure distributions is partially responsible for the wealth of knowledge that is known about the Bell structure. It is also reasonable to speculate that quantum non-classicality in the triangle can be translated directly into a computational advantage for certain computational circuits~\cite{Terhal_2002}. Characterizing quantum non-classicality for interesting causal structures remains an unsolved problem, and the Triangle structure is a small, non-trivial case study.

    Additionally, \citet{Fritz_2012} demonstrated that the Triangle structure is the \textit{smallest} correlation scenario\footnote{Correlation scenarios differ from the Bell structure in that measure setting variables are non-existent in correlation scenarios.} in which their exists incompatible, quantum distributions by explicitly mentioning one which we have elected to call the Fritz distribution (see \cref{sec:fritz_distribution}). Fritz's proof of non-classicality does not involve compatibility inequalities, but instead mentions that inequalities would be helpful for characterizing non-classical in the Triangle structure.
    \comment[E]{Revisions processed up to here thusfar.}

    Finally \citet{Inflation} proposed the Inflation Technique (summarized in \cref{sec:inflation_technique}), a novel tool for solving causal inference problems. Using the inflation technique, they were able to prove incompatibility between the Triangle structure and the $W$-type distribution; a quantum non-accessible distribution whose incompatibility is not witness-able by \textit{any} entropic inequalities or any other known constraints. Therefore as a an algorithmic framework, the Inflation Technique comes closer to characterizing non-classicality in the Triangle structure that all others.

    It is worth noting that a number of causal compatibility inequalities for the Triangle structure have been derived in previous works. For example, \citet{Steudel_2010} derived an inequality for a series of causal structures with $n \in \N$ variables where at most $c \in \N$ share a common ancestor. The Triangle structure is a special case with $c = 2$, $n = 3$. Moreover, \citet{Henson_2014} derive a family of entropic inequalities constraining the set of compatible correlations. Most impressively, \citet{Inflation} derive a family of polynomial, $2$-outcome, causal compatibility inequalities; $37$ non-trivial representatives are printed in~\cite{Inflation}.

    As a preliminary search for quantum incompatibility in the Triangle structure, we performed numerical optimizations using bipartite qubit density matrices and $2$-outcome POVM measurements against the compatibility inequalities printed in~\cite{Inflation} as well as the entropic inequalities presented by \citet{Henson_2014}. Unfortunately, none of these inequalities could be violated. These early results suggest, although do not prove, that quantum non-classical does not exist in the Triangle structure for \textit{two-outcome} measurements. We make \textit{no claim} that this is a universal truth and simply take it as motivation to explore incompatibility in Triangle structure for a degree of outcomes.

    In conclusion, the Triangle structure is a causal structure whose non-classicality remains an open problem. The Triangle structure is a desirable case study because there is a quantum distribution that is known to be incompatible, but no previously known inequalities were able to witness its incompatibility. This failure represents a gap in our understanding of quantum non-classicality, where recent developments (namely the Inflation Technique) have made partial progress. This document expands on our understanding of non-classicality in the Triangle structure and also our understanding on how to find inequalities using the Inflation Technique for interesting causal structures in general.

    \section{Fritz Distribution}
    \label{sec:fritz_distribution}
    \begin{figure}
    \begin{center}
        \scalebox{1.0}{\input{../../figures/causal_structures/triangle_structure_operational_fritz_bell}}
        \caption{The Triangle structure re-imagined to mimic the Bell structure. The measurement settings $S_{\p{A}},S_{\p{B}}$ are latent nodes unlike the Bell structure (\cref{fig:bell_scenario}).}
        \label{fig:triangle_structure_with_fritz_bell_embedded}
    \end{center}
    \end{figure}

    As was originally noticed by \citet{Fritz_2012}, it is possible to construct quantum distributions incompatible with the Triangle structure by utilizing quantum distributions incompatible with the familiar Bell structure. To explain, imagine rearranging the triangle into the configuration depicted in \cref{fig:triangle_structure_with_fritz_bell_embedded} and contrast it with the Bell structure of \cref{fig:bell_scenario}. Evidently, under the correct relabeling, large portions of the Triangle structure resemble the Bell structure. The crucial distinction is that the measurement settings $S_{\p{A}}, S_{\p{B}}$ are \textit{observable variables} in the Bell structure but become \textit{latent variables} in the Triangle structure. In order to embed distributions that are incompatible with the Bell structure into the Triangle structure, the latent nodes $S_{\p{A}}, S_{\p{B}}$ need to be observable and independent of the shared latent node $\la$. This can be accomplished by having $C$ ``measure'' the measurement settings $S_{\p{A}}, S_{\p{B}}$ and announce them as an outcome. Consequently, any distribution over $\p{A}, \p{B}, S_{\p{A}}, S_{\p{B}}$ that is incompatible with Bell structure is also incompatible with the Triangle structure provided that $C$ is \textit{perfectly correlated} with $S_{\p{A}}, S_{\p{B}}$. Understandably there are numerous ways that this can be accomplished, albeit we will focus on elucidating an exemplary method.

    The \term{Fritz distribution}, denoted $\prob[\fritz]$, is a quantum-accessible distribution known to be incompatible with the Triangle structure~\cite{Fritz_2012}. In the Fritz distribution, each of the variables $A,B,C$ have $4$ possible outcomes. Explicitly, $\prob[\fritz]$ can be written as:

@@

    Here the notation $\prob[\fritz][abc] = \prob[\p{ABC}][abc] = \prob[][\p A=a,\p B=b,\p C=c]$ is used as shorthand. The Fritz distribution is best visualized as a $4 \times 4 \times 4$ grid of possible outcomes as depicted in \cref{fig:fritz_distribution_visualized}. From this diagram, it can be seen that each of $\p{C}$'s outcomes restricts the possible outcomes for $\p A,\p B$ into a $2 \times 2$ block. If one writes the outcome labels $\bc{0,1,2,3}$ in binary $\bc{00,01,10,11}$, it can be seen that the left-hand bits for $\p A$ and $\p B$ (respectively denoted $\p A_l$, $\p B_l$) are fixed by the outcome of $\p C$. Pursuant to the embedding of~\cref{fig:triangle_structure_with_fritz_bell_embedded} the left-hand bits emulate the measurement settings $S_{\p A}$ and $S_{\p B}$ and the right-hand bits emulate a two-outcome measurement performed by $A, B$ as they would in~\cref{fig:bell_scenario}. Specifically, $\p C$'s bits are perfectly correlated with the left-bits of $A,B$; $\p C_l = \p A_l$ and $\p C_r = \p B_l$.
